Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-03 BN_discretizer_paper.tex
update

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx} % Allows including images
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[round]{natbib}
\setcitestyle{aysep={ }} %Author-Year SEParator
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Learning Discrete-valued Bayesian Network from Mixed Data%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\author{First Author         \and
        Second Author %etc.
}

@@

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{Discretization \and Bayesian Network \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Bayesian networks \citep{Pearl_1988, PGM_2009}  are an increasingly popular methods for modeling uncertainty and causality in science and engineering. They provide an efficient factorization of the joint probability distribution over a set of random variables. Bayesian networks first emerge from artificial intelligence research and have been applied to a wide variety of problems, ranging from decision-making systems \citep{DMU_2015} to medical diagnoses. In most cases, we assume that all random variables in Bayesian networks are discrete, since many algorithms on Bayesian networks are unable to deal with continuous variables efficiently. However, this assumption is often too restrictive. For example, in the decision-making system of autonomous cars, it is inevitable to deal with continuous variables such as position and velocity.

There are two solutions around this assumption. The first one is to model conditional probability density of each continuous variable by specific families of parametric distributions, then redesign algorithms on Bayesian networks based on these parameters. One successful example is the belief propagation in Gaussian graphical models \citep{Weiss_2011}. Nevertheless, for other shapes of particles \citep{Ihler_2009} or non-linear functions, algorithms are computationally expensive and do not perform well.

The second solution is to discretize continuous variables. Discretization that learns from data has developed well and discussed \citep{Dougherty_1995} in the fields of machine learning and statistics for many year. Most of these discretization methods are designed for classification problems. They search the best discretization of a continuous attribute by considering its interation with the class variable of interest. However, these discretization methods do not apply to continuous variables in Bayesian network. In Bayesian network, interations and dependencies between variables are determined by graph structure. Therefore, if a discretization method only uses the interation between a continuous variable and its class variable of interest, instead of considering the graph structure, then it is not an appropriate discretization method for Bayesian networks. There are some reseach on discretizating continuous variables in naive Bayesian network and tree augmented network \citep{Fried_naive}. Nevertheless, only few discretization methods on general Bayesian netowrk have been proposed \citep{Friedman_1996, Kozlov_1997, Monti_1998}.

The discretization technique proposed by N. Friedman and M. Goldszmidt \citet{Friedman_1996} is the most successful and practical one among these methods. It is based on minimum description length principle (MDL): optimal discretization policy minimizes the description length of Bayesian network and data information. If there is only one continuous variable in Bayesian network and other variables are discrete, MDL discretization method takes running time $O(N^3 + n_c {v_{max}}^{(n_c^p)_{max}} \cdot N^2 + {v_{max}}^{n_p} \cdot N^2)$, where $N$ is the number of data instances for learning the discretization, $n_p$ and $n_c$ are the numbers of parent and children variables for the continuous variable, $v_{max}$ is the largest cardinality number over all variables in Markov blanket, and  ${(n_c^p)_{max}}$ is the largest number of parent variables of the continuous variable's children.

In this paper, we propose a new discretization method for continuous variables in Bayesian network by learning from mixed data. This method is a generalized version of  \citet{Boulle_2006} and \citet{Lustgarten_2011}, which are both discretization methods of a continuous attribute with one class variable. We begin our method with the situation that only one variable in a given Bayesian network is continuous and other variable are discrete. Under this situation, we look for the most possible discretization policy $M$ given the data on other discrete variables. That is to say, the desired policy is $\textrm{arg}_M P(M|D)$. With Bayes rule, $P(M|D) = P(M) \cdot P(D|M) / P(D) \propto P(M)\cdot P(D|M)$. Usually, $P(D|M)$ increases as number of discretized intervals increases, since more intervals can provide more accuracy. Therefore, we design probability priors of $P(M)$ to decrease as number of intervals rises. As a result, we obtain a nature trade-off to determine the number of intervals after discretization. Besides, the $P(D|M)$ can be factorized according to Bayesian network structure. On the other hand, with the proposed priors, we are able to restrict the number of discretized intervals not exceeding the largest cardinality number of variable in Markov blanket too much. This is important, since for most algorithms on Bayesian network, their running times exponentially depend on the cardinality of variables. Another advantage of our algorithm is running time. If there is only one continuous variable, the running time is $O(n_c  {v_{max}} \cdot N^2 + {v_{max}}^{n_p} \cdot N^2)$, which beats the running time of MDL principle discretization \citep{Friedman_1996}. With an effective approximation proposed in this paper, we can further reduce running time to $O({v_{max}}(n_p+n_c)N^2)$.

For a Bayesian network with multiple continuous variables, we apply the discretization method discussed above iteratively. In the beginning, we discretize each continuous variable by equal-width method. Number of intervals for equal-width discretization is equal to largest cardinality of discrete variables. Then we iterate the one-variable discretization on each continuous variable by the following order: from the variable with highest topological order (leaves) to the variable with lowest topological order (root). Everytime when we finish a discretization procedure on one variable, we store the discretization result and this discretization result will be used for later discretization on another variable. Experiments on real data show that with same iteration order, our discretization provides better discretization result than MDL principle method in terms of likelihood. The latter is easily catched by local minima and thus produce non-necessarily low number of intervals after discretization.

Finally, we can combine the new discretization method with K2 structure learning alorithm \citep{K2}. We first prediscretize all continuous variables before K2 learning, since K2 algorithm requires all variables are discrete. Everytime when an edge is added by K2 algorithm procedure, we rediscretize the relevant variables in Bayesian network and store the result for next K2 algorithm iteration. By this principle, we are able to learn a discrete-valued Bayesian network from mixed data.

The rest of the paper is organized as follows: we prepare the preliminaries of Bayesian network in Section 2. Related work, including MDL principle discretization \citep{Friedman_1996} and MODL discretization by \citet{Boulle_2006}, are summarized in Section 3. In Section 4, we introduce the new discretization method, including formulation of objective function and algorithms. One continuous variable case and multiple continuous variable case are also included in this section. Finally, in Section 5, we apply the new discretization method to real-world data and show the result.

\section{Preliminaries}
\label{Prelim}
\subsection{Bayesian Network}
\label{BN_basic}
A Bayesian network $B$ is defined by a pair $(G,\Theta)$, where $G = (X,E)$ is a directed acyclic graph whose nodes correspond to a set of random variable $X = \{  X_1, X_2, \cdots, X_n \}$, and whose edges $E$ represents probabilistic dependencies among nodes. The graph structure $G$ encodes the Markov property: each node $X_i$ is independent of its non-descendants given its parents in $G$. The second component of $B$, namely $\Theta$, contains a set of parameter that qualifies the network. Elements of $\Theta$ take the form $\theta_{x_i|\Pi_{x_i}} = P(x_i|\Pi_{x_i})$ for each possible value $x_i$ of $X_i$, and $\Pi_{x_i}$ of $\Pi_{X_i}$ (the set of parents of $X_i$ in $G$). Due to Markov property, we can represent the multivariate joint distribution over $X$ as
\begin{equation*}
P_B (X_1 , \cdots, X_n) = \prod_{i=1}^{n} P_B (X_i | \Pi_{X_i}) = \prod_{i=1}^{n} \theta_{x_i|\Pi_{x_i}}
\end{equation*}\\
For a given Bayesian network, we can generate its data instances by forward sampling, i.e., sampling variables one by one in a topological order \citep[see][chap.~22]{algo_2009}. Given a instance of $X_i$'s parent set $\Pi_{X_i}$, value of $X_i$ can be sampled according to the conditional probability table $P(X_i | \Pi_{X_i})$. Furthermore, this parent-child sampling order can be reversed if we know the marginal probability of child variable $P(X_i)$. By Bayes rule,
\begin{equation*}
P(X_i | \Pi_{X_i}) \cdot \prod_{j = 1}^{ Pa( X_i)} P( \{ \Pi_{X_i} \}_j) = P(X_i) \cdot P(\Pi_{X_i} | X_i),
\end{equation*}
where $Pa( X_i)$ is the number of parents of $X_i$, and $\{ \Pi_{X_i} \}_j$ is the $j$th parent of $X_i$. We can first sample $X_i$ by $P(X_i)$, then sample all the parent varaibles simultaneously by $P(\Pi_{X_i} | X_i)$. That is to say, $X_i$ becomes the starting point for sampling.

\subsection{K2 Structure Learning}
\label{intro_K2}
K2 ...

\section{Related Work}
\label{relat_work}
\subsection{MDL Principle Discretization}
\label{MDL}

\subsection{MODL Discretization}
\label{MODL}

\section{Discretize One Continuous Variable}
\label{one-conti}

\subsection{Notations}
\label{Notation}

\subsection{Objective Function}
\label{obj}

\subsection{Algorithm}
\label{algo}

\subsection{Approximation}
\label{approx}

\section{Discretizing Multiple Continuous Variable with Structure Learning}
\label{multiple-conti}

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx} % Allows including images
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[round]{natbib}
\setcitestyle{aysep={ }} %Author-Year SEParator
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Learning Discrete-valued Bayesian Networks from Mixed Data%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

\author{First Author         \and
        Second Author         \and
        Third Author%etc.
}

@@

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{Discretization \and Bayesian Network \and Continuous Variable}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Bayesian networks \citep{Pearl_1988, PGM_2009}  are an increasingly popular method for modeling uncertainty and causality in science and engineering. They provide an efficient factorization of the joint probability distribution over a set of random variables. Bayesian networks first emerged from artificial intelligence research and have been applied to a wide variety of problems, ranging from decision-making systems \citep{DMU_2015} to medical diagnoses. In most cases, we assume that all random variables in Bayesian networks are discrete, since many algorithms on Bayesian networks are unable to deal with continuous variables efficiently. However, the assumption that all variables are discrete is often too restrictive. For example, in the decision-making system of autonomous cars, it is imperative to deal with continuous variables such as position and velocity.

There are two methods around this assumption. The first one is to model conditional probability density of each continuous variable by specific families of parametric distributions, then redesign algorithms on Bayesian networks based on these parameters. One successful example is belief propagation in Gaussian graphical models \citep{Weiss_2011}. Nevertheless, for other shapes of particles \citep{Ihler_2009} or non-linear functions, these redesigned algorithms are computationally expensive and do not perform well.

The second method around continuous variables is to discretize them. Discretization that learns from data has been developed well and discussed in the fields of machine learning and statistics for many years \citep{Dougherty_1995, Kerber_1992, Holte_1993, Fayyad_1993} . Most of these discretization methods are designed for classification problems. They search the best discretization policy of a continuous attribute by considering its interation with the class variable of interest. However, these discretization methods do not apply to continuous variables in Bayesian networks. In Bayesian networks, interations and dependencies between variables are determined by graph structure. Therefore, if a method discretizes continuous variables according to graph structure, instead of assigning one variable as a class variable, then it would be more appropriate. There is some reseach on discretizing continuous variables in naive Bayesian networks and tree augmented networks \citep{Fried_naive}, but only a few discretization methods on general Bayesian netowrk have been proposed \citep{Friedman_1996, Kozlov_1997, Monti_1998, Steck_2007}.

The discretization technique proposed by \citet{Friedman_1996} is the most well-known one among these few methods. It is based on minimum description length principle (MDL): optimal discretization policy minimizes the description length of Bayesian network and data information. If there is only one continuous variable in Bayesian network and other variables are discrete, this MDL discretization method takes running time $O(N^3 + n_c {v_{max}}^{(n_c^p)_{max}} \cdot N^2 + {v_{max}}^{n_p} \cdot N^2)$, where $N$ is the number of data instances for learning the discretization, $n_p$ and $n_c$ are the numbers of parent and children variables for the continuous variable, $v_{max}$ is the largest cardinality number over all variables in Markov blanket, and  ${(n_c^p)_{max}}$ is the largest number of parent variables of the continuous variable's children. If there are multiple continuous variables in the network, the method iterates over all continuous variables until convergence. In our test on real world data, the iterations converge in a few cycles. However, the number of intervals after discretization is too small.

In this paper, we propose a new discretization method for continuous variables in Bayesian networks by learning from mixed data. This method is a generalized version of  \citet{Boulle_2006} and \citet{Lustgarten_2011}, which are both discretization methods of a continuous attribute with one class variable. We begin our method with the assumption that only one variable in a given Bayesian network is continuous and other variables are discrete. Furthermore we assume that the network structure is known in advance. Under this situation, we look for the most probable discretization policy $M$ given the data $D$ on other discrete variables. That is to say, the desired policy is $\textrm{arg}_M P(M \mid D)$. With Bayes' rule, $P(M \mid D) $ can be rewritten as $P(M) \cdot P(D \mid M) / P(D)$ , which is proportional to $P(M)\cdot P(D \mid M)$. Usually, $P(D \mid M)$ increases as the number of discretized intervals increases, since more intervals can provide more accuracy. $P(M)$, on the other hand, is designed to decrease as the number of intervals rises. As a result, we can automatically determine the number of intervals after discretization by maximizing  $P(M)\cdot P(D \mid M)$. In addition, the $P(D \mid M)$ can be factorized according to Bayesian network structure. With the proposed priors, we are able to restrict the number of discretized intervals so that it will not exceed the largest cardinality number of variables in the Markov blanket by too much. This is important, since for most algorithms on Bayesian networks, their running times exponentially depend on the cardinality of variables. Another advantage of our method is the running time for learning. If there is only one continuous variable, the running time is $O(n_c  {v_{max}} \cdot N^2 + {v_{max}}^{n_p} \cdot N^2)$, which is significiantly shorter than the running time of MDL principle discretization \citep{Friedman_1996}. With the approximation proposed in this paper, we can further reduce running time to $O({v_{max}}(n_p+n_c)N^2)$.

For a Bayesian network with multiple continuous variables, we apply the discretization method discussed above iteratively. In the beginning, we prediscretize each continuous variable by equal-width method. The number of intervals for equal-width discretization is equal to largest cardinality of discrete variables. Then we iterate the one-variable discretization on each continuous variable in the following order: from the variable with highest topological order (leaves) to the variable with lowest topological order (root). Each time we finish a discretization procedure on one variable, we store the discretization result for later iterations. Experiments on real data show that with the same iteration order, our discretization method provides better discretization results than MDL principle method in terms of likelihood. The MDL principle method is easily stuck at local minima and also discretizes variables into too few intervals.

Finally, we can combine the our new discretization method with K2 structure learning alorithm \citep{K2}. We first prediscretize all continuous variables before K2 learning, since the K2 algorithm requires all variables to be discrete. Each time an edge is added in the K2 algorithm procedure, we rediscretize the relevant variables in the learned Bayesian network and store the result for next K2 algorithm iteration. By this principle, we are able to learn a discrete-valued Bayesian network from mixed data.

The rest of the paper is organized as follows: related works, including MDL principle discretization \citep{Friedman_1996} and MODL discretization by \citet{Boulle_2006}, are summarized in Section 2. We prepare the preliminaries and notations in Section 3. In Section 4, we introduce the new discretization method for the case that only one variable is continuous, including the derivation of objective function and the algorithms. Section 5 is a discussion about multiple continuous variables in a Bayesian network and how to discretize them. The combination of our proposed discretization and K2 structure learning is also introduced. Finally, in Section 6, we apply the proposed discretization method to real-world data and show the result.

\section{Related Work}
\label{relat_work}
In this section we review two related works: MDL principle discretization \citep{Friedman_1996} and MODL discretization \citep{Boulle_2006}. The former is the most famous discretization method for continuous variables in Bayesian networks, and we will compare it with our proposed method in Section 6. The latter is a discretization method for one continuous attribute according to a target class. Our proposed method is a generalization of this method. Note that MODL discretization is a Bayesian approach. The asymptotical equivalence between MDL approach and Bayesian approach has been examined in \citep{VL_2000}.
\subsection{MDL Principle Discretization}
\label{MDL}
The MDL principle was first proposed by \cite{MDL_1978}. It provides a way to compress data, that is to say, describe data with fewer number of symbols than the number of symbols needed to describe the data literally \citep{Grunwald_2009}. MDL chooses a model that trades off goodness-of-fit for the complexity of model. Therefore it has an advantage that automatically avoids overfitting. In the case of Bayesian networks, \citep{Friedman_1996} applied the MDL concept to determine the number of intervals of each continuous variable after discretization and also the positions of discretization boundaries. Here is the mechanism: MDL principle discretization selects a discretization policy that minimizes sum of description lengths of discretized Bayesian network and the necessary information that recovers original data from discretized data. If there is only one continuous variable and other variables are discrete, the objective function is  
\begin{equation}
\begin{aligned}
&{1 \over 2} log(N) \{  \lVert \Pi_{X_i} \rVert (\lVert X^*_i \rVert - 1) +
 {\sum_{j,X_i \in \Pi_{X_i}}} \lVert \Pi_{X^*_j} \rVert (\lVert X_j \rVert - 1) \} + \\
& log(\lVert {X^*_i} \rVert) + log{{N_{i}-1}\choose{\lVert {X^*_i} \rVert -1}}  -N \cdot [I(X^*_i,\Pi_{X_i}) + {\sum_{j,X_i \in \Pi_{X_i}}} I(X_j, \Pi_{X^*_j})],
\end{aligned}
\end{equation}
where function $I(\boldsymbol{A},\boldsymbol{B}) = \sum_{\boldsymbol{a},\boldsymbol{b}} \hat{P}_D (\boldsymbol{a},\boldsymbol{b}) log {{\hat{P}_D (\boldsymbol{a},\boldsymbol{b})}\over{\hat{P}_D (\boldsymbol{a}),\hat{P}_D (\boldsymbol{b})}}$ is the mutual information between two sets of variables $\boldsymbol{A}$ and $\boldsymbol{B}$, $||\boldsymbol{X}||$ means the cardinality of variable $\boldsymbol{X}$ and $X^*_i$ is the discretized version of continuous variable $X_i$. The optimal discretization policy can be found by dynamical programming. If there are multiple continuous variables in Bayesian network, we apply the algorithm on one variable at a time and iterate over all continuous variables. While iterating, only one variable is treated as continuous and other continuous variables are discretized based on initial prediscretization or the discretization result of previous iteration.

For the above discretization processes, the network structure is known in advance. If the network structure is not given, we can alternate between structure learning and discretization learning. We start with some discretization, and learn a network structure given this discretization. Then, we rediscretize based on the learned network. This cycle continues until no improvement in objective function is made.
\subsection{MODL Discretization}
\label{MODL}
MODL discretization \cite{Boulle_2006} is a Bayesian-approach discretization. It discretizes a continuous feature according to a class variable. The best discretization model is found by maximizing the probability $P(\textit{Model} \mid \textit{Data})$ of the model given the data. Using the Bayesian rule and since $P(\textit{Data})$ is a constant while varying the discretization model, this is equivalent to maximize:
\begin{align}
P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model}).
\end{align}
Once the prior distribution of the discretization model is fixed, each term in Eq (3) can be evaluated. Usually, $P(\textit{Data} \mid \textit{Model})$ increases as the number of intervals of discretization model increases, since more intervals means that the model contains more information and thus it is easier to reproduce the original data. On the other hand, $P(\textit{Model})$ decreases as the number of intervals increases. Thus, maximizing $P(\textit{Model}) \cdot P(\textit{Data} \mid \textit{Model})$ provides a trade-off to determine number of intervals after discretization.

The MODL method adapts dynamical programming to find the optimal discretization model, and it takes running time $O(N^3 + ||\text{Class}|| \cdot N^2)$, where $N$ is number of data instances. \cite{Lustgarten_2011} provide a different prior to evaluate $P(\text{Model})$ term. This prior could include the assumption of uniform prior probabilities over discretization in MODL as a special case. Additionally, it reduces runing time to $O(||\text{Class}|| \cdot N^2)$.

\section{Preliminaries}
\label{Prelim_Notation}
In this section we provide a brief review of Bayesian networks, including the factorization of joint probability distribution, sampling from a given network, and structure learning. These concepts will be used in later sections.

\subsection{Bayesian Network and Structure Learning}
\label{BN_basic}
A Bayesian network $B$ is defined by a pair $(G,\Theta)$, where $G = (X,E)$ is a directed acyclic graph whose nodes correspond to a set of random variable $X = \{  X_1, X_2, \cdots, X_n \}$, and whose edges $E$ represents probabilistic dependencies among nodes. The graph structure $G$ encodes the Markov property: each node $X_i$ is independent of its non-descendants given its parents in $G$. The second component of $B$, namely $\Theta$, contains a set of parameters that qualify the network. Elements of $\Theta$ take the form $\theta_{x_i \mid \Pi_{x_i}} = P(x_i \mid \Pi_{x_i})$ for each possible value $x_i$ of $X_i$, and $\Pi_{x_i}$ of $\Pi_{X_i}$ (the set of parents of $X_i$ in $G$). Applying Markov property, we can represent the multivariate joint distribution over $X$ as
\begin{equation}
P_B (X_1 , \cdots, X_n) = \prod_{i=1}^{n} P_B (X_i \mid \Pi_{X_i}) = \prod_{i=1}^{n} \theta_{x_i \mid \Pi_{x_i}}
\end{equation}

For a given Bayesian network, we can generate its data instances by forward sampling, i.e., sampling variables one by one in a topological order \citep[see][chap.~22]{algo_2009}. Given an instance of $X_i$'s parent set, $\Pi_{X_i}$, values of $X_i$ can be sampled according to the conditional probability table $P(X_i \mid \Pi_{X_i})$. Furthermore, this parent-child sampling order can be reversed if we know the marginal probability of child variable $P(X_i)$. By Bayes rule,
\begin{equation}
P(X_i \mid \Pi_{X_i}) \cdot \prod_{j = 1}^{ Pa( X_i)} P( \{ \Pi_{X_i} \}_j) = P(X_i) \cdot P(\Pi_{X_i} \mid X_i),
\end{equation}
where $Pa( X_i)$ is the number of parents of $X_i$, and $\{ \Pi_{X_i} \}_j$ is the $j$th parent of $X_i$. We can first sample $X_i$ by $P(X_i)$, then sample all the parent variables simultaneously by $P(\Pi_{X_i} \mid X_i)$. That is to say, $X_i$ becomes the starting point for sampling.

If we don't know the structure of the Bayesian network in advance, then we need to learn it from data. Roughly speaking, there are three approaches to learn a Bayesian network structure from data \citep[see][chap.~18]{PGM_2009}: constraint-based structure learning, score-based structure learning, and Bayesian model averaging. Here we review K2 structure learning algorithm \citep{K2}, which is one of the most successful score-based structure learning methods. Similar to most structure learning algorithms, the K2 algorithm requires all variables to be discrete. The score of a learned network is defined as $\prod_i f(X_i, \Pi_{X_i})$, where
\begin{align}
f(X_i, \Pi_{X_i}) = \prod_{j=1}^{||\Pi_{X_i}||} {{(||X_i||-1)!}\over{(N_{ij} + ||X_i||-1)!}} \prod_{k=1}^{||X_i||} \alpha_{ijk}!.
\end{align}
$||X_i||$ is the cardinality of variable $X_i$. $||\Pi_{X_i}||$ is number of all possible instantiations of the parent variables of $X_i$, i.e., $|\Pi_{X_i}| = \prod_{Y \in \Pi_{X_i}} |Y|$. $\alpha_{ijk}$ is the number of instances in a data set that variable $X_i$ is instantiated with its $k$th value, and the parent of $X_i$ are instantiated with the $j$th value of $\Pi_{X_i}$. $N_{ij} = \sum_{k=1}^{|X_i|} \alpha_{ijk}$.

K2 algorithm searches the network structure with the highest score, which can be interpreted as the most probable network with the given data. In addition, K2 requires a topological order of variables to be known before scoring can start. This constraint can prevent cycles from being introduced. The searching for high-score network is an iterative process. There is no way to find the optimal network directly, since there are $2^{O(n^2)}$ possible structures, where $n$ is number of variables. In order to compensate for this, we run K2 algorithm many times, and each time we start with a different order of variables. The network with the highest score in iterations is the desired one.

\subsection{Discretization Policy}
\label{Disc_p}
A Discretization Policy $M_C = < t_1,t_2, ..., t_{k-1} >$ on a continuous variable $C$ is a mapping from $\boldsymbol{R}$ to $\{1,2,3,...,k \}$ such that
\begin{equation}
  M_C (x)=\begin{cases}
    1, & \text{if $x<t_1$}.\\
    i, & \text{if $t_{i-1} \leq x < t_i$}.\\
    k, & \text{if $t_k \leq x$}
  \end{cases}
\end{equation}
That is to say, the policy discretizes the continuous variable $C$ into $k$ intervals. Furthermore, we assume that the discretization edge $t_i$ can only take values on middle points of two consective values of data of $C$. By this assumption, we can obtain an integer representation of $M_C$ as follows: 
we sort the data of continuous variable $C$ in an increasing order, $\{ c_1,c_2,...,c_N \}$, and if $t_i = (c_{s_{i}} + c_{s_{i}+1})/2$ for $i=1,2,...,k$, then $M_C = < t_1,t_2, ..., t_{k-1} > \equiv [n_1,n_2,..n_k]$, where $n_1 = s_1$ and $n_i = s_{i} - s_{i-1}$ for $i = 2,...,k$.\\

\section{Discretize One Continuous Variable}
\label{one-conti}
In this section, we consider the case that only one variable in a Bayesian network is continuous and the others are discrete. Before we formulate the objective function, we introduce some notations that will make the following calculation easier.
\subsection{Notations}
\label{Notation}
Assume the continuous variable is $X$. It has $n_p$ parent variables, $\Pi_X = \{ P_1, P_2, P_3,...,P_{n_p}\}$. It also has $n_c$ child variables $\{ C_1, C_2,...,C_{n_c}\}$. Each child variable $C_i$ of $X$ has other parent variables, which forms a set $\boldsymbol{S_i}$. Before we do the discretization, we sort the data instances by its attribute of $X$.

Let $D \cup D_X$ be a dataset of $N$ instances which we plan to learn discretization from. $D_X$ only contains the data of $X$. $D$ contains the data of all variables except $X$.  Assume $D \cup D_X$ has been sorted in the ascending order of $D_X$. In the following content, $D_{\boldsymbol{Y}}$ means the data instances of a set of variable $\boldsymbol{Y}$. Let $D_X = \{ x_1,x_2,x_3,...,x_N \}$. In order to make the discretization mechanism more understandable, we temporarily assume there is no repeated values in $D_X$, i.e., $x_1 < x_2 < ... < x_N$. We will remove this assumption later in the algorithm section. Therefore, a discretization policy $M$ on $X$ can be written as $M = [n_1,n_2,...,n_k]$, where $k$, $n_1$, $n_2$,..., and $n_k$ are positive integers satisfying $N = \sum_{i=1}^k n_i$.

%Furthermore, since there might have repeated values for $D_x$, we assume there are $N'$ unique values of $D_x$ and $N' \leq N$. Note that the allowable discretization edges happens between two consective values of non-repeated value of $D_x$. For example, if $D_X = \{ 1.0,1.0,2.0\}$, then the only possible discretization edge is on $1.5$.

\subsection{Priors and Objective Function}
\label{obj}
We have the following four priors for the discretization model $M$ that enable us to evaluate $P(M)$ and $P(D \mid M)$:\\
\begin{enumerate}
\item For a discretization edge locating at $(x_i + x_{i+1})/2$ has probability
\begin{equation}
1 - {exp(- M \cdot {{x_{i+1} - x_i}\over{x_N - x_1}})},
\end{equation}
where $M$ is the largest cardinality number of discrete variables in $X$'s Markov blanket.
\item For a given interval of $M$, every distribution of $\Pi_X$ is equiprobable.
\item For each pair of $(C_i,\boldsymbol{S_i})$ and a given interval of $M$, every distribution of $C_i$ with a value of $\boldsymbol{S_i}$ is equiprobable.
\item The distributions of each $C_i$ and $\Pi_X$ in each interval of $M$ are independent from each other.
\end{enumerate}

Owing to these priors, we are able to evaluate $P(M)$ and $P(D \mid M)$. For the former, we have
\begin{equation}
P(M) = \prod_{i=1}^k \{ [1 - {exp(- M \cdot {{x_{s_i+1} - c_{x_i}}\over{x_N - x_1}})}] \cdot exp(M \cdot {{x_{s_{i}} - x_{s_{i-1} + 1}}\over{x_N - x_1}}) \},
\end{equation}
where $s_i = \sum_{j=1}^k n_j$ and $s_k = N$, $s_0 = 0$. Before evaluating $P(D \mid M)$, notice that
\begin{equation}
P(D \mid M) \propto P(D_{\Pi_x} \mid M) \cdot \prod_{i = 1}^{n_c} P(D_{C_i} \mid M, D_{\boldsymbol{S_i}}).
\end{equation}
The graph structure allows us to only consider the interations between $X$ and other variables in its Markov blanket and also make the factorization. For example, in Figure 1., the corresponding $P(D \mid M)$ is factorized as
\begin{equation}
P(D_{ P_1,P_2,P_3 } \mid M) \cdot P( D_{ C_1 } \mid M,D_{\boldsymbol{S_1}}) \cdot P(D_{C_2} \mid M,D_{ S_2  }).
\end{equation}

% For one-column wide figures use
\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[width=6cm]{graph1.png}
% figure caption is below the figure
\caption{Factorization of $P(D \mid M)$}
\label{fig:1}       % Give a unique label
\end{figure}

The concept behind the factorization is similar to sampling in a Bayesian network. Once we have the distribution of $X$ and sample $X$ for it, then we can further sample $\Pi_x$. Since the samples of variables in $\Pi_x$ are not independent, we can not further factorize $P(D_{\Pi_x} \mid M)$ into $\prod_i P(D_{P_i} \mid M)$. Similarly, once we have a set of samples of $X$ and $\boldsymbol{S_i}$, we can sample $C_i$. Since the sampling processes of child variables are independent from each other, we have the product $\prod_{i = 1}^{n_c} P(D_{C_i} \mid M, D_{\boldsymbol{S_i}})$.\\
